---
title: "Lesson_1_test"
author: "Yuying"
date: "August 29, 2022"
output: 
  html_document: 
    toc: yes
---

## Basic commands

Example of some basic math operations
```{r}
sqrt(25)
4+8
5*8
```

## Basic plot
```{r}
x<-1:19
plot(x)
```

## Data types

### Numeric
```{r}
# concatenate
xdata<-c(3,7,4,3)
xdata
mean(xdata)
```

### Character

```{r}
names<-c("sherry","susie","iris","this is r course")
names
```

### Factor

```{r}
# non norminal
fact1<-c("low","low","high","moderate","high","moderate")
fact1
# get frequencies
table(fact1)
barplot(table(fact1))

```

### nominal

```{r}

nomfac<-c("m","f","x","x","m","f","x","x")
nomfac
nomtab<-table(nomfac)
barplot(nomtab)
```

### logical

```{r}
ans<-c("T","F","T","F","T","F","T","F")
ans
anstab<-table(ans)
barplot(anstab)
```

### variable type

```{r}
class(ans)
class(xdata)
```

### convert between data types

```{r}
factfact1<-as.factor(fact1)
factfact1
```

### exercise

```{r}
ex1<-c(1,"two",3,"four")
ex2<-c(3.14,5,8,1.2)
ex3<-c("Belgium","Spain","France")
ex4<-c("true","false","true")
ex5<-c(F,F,F,T)
class(ex1)
class(ex2)
class(ex3)
class(ex4)
class(ex5)
```

## Data structures

### constant

```{r}
a<-7
b<-14.5
c<--5
a
b
c
```

### vectors

#### c()
```{r}
xvec<-c(4,3,5,7)
xvec
```

#### index
```{r}
xvec2<-1:10
xvec2
```

#### rep
```{r}
xvec3<-rep(1:5,3)
xvec3
```

#### seq
```{r}
xvec4<-seq(0,5,0.5)
xvec4
```

#### random number generation
```{r}
# if you want everytime same output set.seed(2022)
rdataunif<-runif(5)
rdataunif

## benomial
radatbin<-rbinom(10,1,0.5)
radatbin

##binomial
radatbin2<-rbinom(10,5,0.4)
radatbin2

## standard normal
rdatanorm<-rnorm(30)
rdatanorm
hist(rdatanorm)

## standard normal
rdatanorm1<-rnorm(30,3,4)
rdatanorm1
hist(rdatanorm1)
```
#### Matrix

```{r}
xmat1<-matrix(1:12, nrow=3,byrow=T)
xmat1

xmat2<-matrix(1:12, nrow=4)
xmat2
```

#### dataframe
```{r}
x1<-c("sherry","susie","iris","kent")
y1<-c(1,3,5,6)
mydata<-data.frame(x1,y1)
mydata

str(mydata)

##matrix to dataframe
matrixttodf<-as.data.frame(xmat1)
matrixttodf
names(matrixttodf)<-c("sherry","iris","susie","kent")
matrixttodf
```
### lists
```{r}
xlist<-list(x1,y1,matrixttodf)
xlist[[3]]
```

### exercise
```{r}
set.seed(258)
test<-runif(10)
test
test[3]
test<-test[10:1]
test
test[3]
test[3:6]
## drop columns 3 to 6
test[-c(3:6)]

test1<-c(1,2,3,4,5,6,7)
test1
test1>5

## quick check
X = 4
Y = 6

X >= 0 & Y < 10
X != 4
Y == 6
X+Y >0 | Y-X < 0

k=c("b","g","r","d")
k[k=="r"]

k[k!="r"]
```

## Functions

## Packages
```{r}
#install.packages("graphics")
library(graphics)
```


## Data import/export

### R built-in data
```{r}
str(cars)
names(cars)
hist(cars$speed)
```

### tab-delimated

```{r}
origin<-read.delim("origin.txt")

## > origin <- read.delim("C:/Users/YTan/Desktop/R exercise/origin.txt", comment.char="#")
```

### csv
```{r}
data<-read.csv("Data.csv")
head(data)
```

### Excel file
```{r}
library(readxl)
xlData <- read_excel("Data.xlsx")
head(xlData)

## use xlsx paclage
#library(xlsx)
#newdata<-read.xlsx("Data.xlsx",1)
#head(newdata)
```

### Exporting data
```{r}
write.table(xlData,"exportedxlfile.txt")
## write.xlsx(xlData,"exportedxlfile.txt")
library(readr)
write_csv(xlData,"exportedxlfile.csv")
```

## Data manipulation

### subsetting
```{r}
library(readr)
Data<-read.csv("Data.csv")
head(Data)

#subset
dtafem<-subset(Data,gender=="female")
head(dtafem)

dtafem2<-subset(Data,gender=="female"&exam<=14)
head(dtafem2)

dtafem3<-subset(Data,gender=="female"|exam>10)
dtafem3
```

## Merge two datasets
```{r}
setos_Sepal<-subset(iris,Species=="setosa",select=c("Sepal.Length","Sepal.Width"))
setos_Sepal$ID<-paste0(1:50,"id")

setos_Petal<-subset(iris,Species=="setosa",select=c("Petal.Length","Petal.Width"))
setos_Petal$ID<-paste0(1:50,"id")

## merge two files
seto<-merge(setos_Sepal,setos_Petal,by="ID")
head(seto)
```

